Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-23 thesis/paper_draft.qmd
revert screeplot back to cov

Many real-world hierarchical data sets may exhibit a prominent low-rank (factor) structure. A clear example is the Australian domestic overnight trips [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions. The tourism activities might be driven by a few common factors, such as economic conditions, fuel prices, or major events, which might be left in the forecast errors after fitting the base models. @fig-eigen illustrates the largest eigenvalues from the one-step forecast error correlation matrix $\hat{\boldsymbol{R}}_1$, typically showing a marked elbow after a handful principal components (largest eigenvalues), indicating strong latent structure. Diagonal-target shrinkage is factor-unaware: it neither preserves the low-rank common subspace nor differentially shrinks the idiosyncratic remainder, and so can be inefficient when common factors dominate.

@@

```{r}
#| label: fig-eigen
#| fig.cap: "Scree plot of the eigenvalues of the 1-step-ahead forecast error covariance matrix from the Australian domestic tourism dataset. The point of inflection occurs around the fourth largest eigenvalue."
#| fig.width: 7
#| fig.height: 4
#| out.width: "70%"
tourism_resid <- readRDS(paste0(path_lib, "tourism_resid.rds"))

@@

(eigen(cor(tourism_resid))$values[1:20]/1e6) %>%
  plot(type = "b",
       xlab = "Component",
       ylab = "Eigenvalue ('000,000)"
  )
# add a vertical line at 5
abline(v = 4, col = "purple", lty = 2, lwd = 2)
```

>>>>

Many real-world hierarchical data sets may exhibit a prominent low-rank (factor) structure. A clear example is the Australian domestic overnight trips [@tourism-au], where the national trips are disaggregated into states and territories, and further into regions. The tourism activities might be driven by a few common factors, such as economic conditions, fuel prices, or major events, which might be left in the forecast errors after fitting the base models. @fig-eigen illustrates the largest eigenvalues from the one-step forecast error correlation matrix $\hat{\boldsymbol{R}}_1$, typically showing a marked elbow after a few principal components (largest eigenvalues), indicating strong latent structure. Diagonal-target shrinkage is factor-unaware: it neither preserves the low-rank common subspace nor differentially shrinks the idiosyncratic remainder, and so can be inefficient when common factors dominate.

@@

```{r}
#| label: fig-eigen
#| fig.cap: "Scree plot of the eigenvalues of the 1-step-ahead forecast error covariance matrix from the Australian domestic tourism dataset. The point of inflection occurs around the second largest eigenvalue."
#| fig.width: 7
#| fig.height: 4
#| out.width: "70%"
tourism_resid <- readRDS(paste0(path_lib, "tourism_resid.rds"))

@@

(eigen(cov(tourism_resid))$values[1:20]/1e6) %>%
  plot(type = "b",
       xlab = "Component",
       ylab = "Eigenvalue ('000,000)"
  )
# add a vertical line at 5
abline(v = 2, col = "purple", lty = 2, lwd = 2)
```
